Internship Summary: Mid-Term Review

### 2020-07-23 RapportStage/Rapport.tex
update sec 5.1

\begin{table}[h!]
    \centering
    \begin{tabular}{c c c c c c c}
\hline
 $r$ & UHF & UMP2 & UMP3 & UMP4 & $\expval{S^2}$ \\
\hline
0.75 & 0.0\% & 63.8\% & 87.4\% & 95.9\% & 0.00\\
1.35 & 0.0\% & 15.2\% & 26.1\% & 34.9\% & 0.49\\
2.00 & 0.0\% & 01.0\% & 01.8\% & 02.6\% & 0.95\\
2.50 & 0.0\% & 00.1\% & 00.3\% & 00.4\% & 0.99\\
\hline
\end{tabular}
    \caption{\centering Percentage of electron correlation energy recovered and $\expval{S^2}$ for the \ce{H2} molecule as a function of bond length (r,\si{\angstrom}) in the STO-3G basis set (taken from \cite{Gill_1988}).}
    \label{tab:SpinContamination}
\end{table}

@@

%============================================================%
\section{The spherium model}
%============================================================%

@@

\begin{table}[h!]
\centering
\begin{tabular}{c c c c c c c c c}
$R$ & 0.1 & 1 & 2 & 3 & 5 & 10 & 100 & 1000 \\
\hline
RHF & 10 & 1 & 0.5 & 0.3333 & 0.2 & 0.1 & 0.01 & 0.001 \\
\hline
UHF & 10 & 1 & 0.490699 & 0.308532 & 0.170833 & 0.078497 & 0.007112 & 0.000703 \\
\hline
Exact & 9.783874 & 0.852781 & 0.391959 & 0.247898 & 0.139471 & 0.064525 & 0.005487 & 0.000515 \\
\end{tabular}
    \caption{\centering RHF and UHF energies in the minimal basis and exact energies for various R.}
\label{tab:ERHFvsEUHF}
\end{table}

@@

The Hamiltonian $\bH(\lambda)$ is block diagonal because of the symmetry of the basis set ($\psi_1$ interacts with $\psi_4$ and $\psi_2$ with $\psi_3$). The two singly excited states yields a singlet and a triplet sp\textsubscript{z} but they don't have the same symmetry so these states can't form exceptional points with the ground state. However there is an avoided crossing (see Fig. \ref{fig:RHFMiniBas}) between the s\textsuperscript{2} and the p\textsubscript{z}\textsuperscript{2} which is connected to two exceptional points in the complex plane.

@@

Then using this basis set we can compare the different partitioning of \autoref{sec:AlterPart}. The \autoref{fig:RadiusPartitioning} shows the evolution of the radius of convergence $R_{\text{CV}}$ in function of $R$ for the M{\o}ller-Plesset, the Epstein-Nesbet, the Weak Correlation and the Strong Coupling partitioning in a minimal basis i.e. $P_0$ and $P_1$. We see that

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\textwidth]{PartitioningRCV.pdf}
    \caption{\centering Radius of convergence in the minimal basis for different partitioning of the Hamiltonian$\bH(\lambda)$.}
    \label{fig:RadiusPartitioning}
\end{figure}

Commenter figure 8 (singularité alpha pas très sensible à basis set)

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.45\textwidth]{MPlargebasis.pdf}
    \includegraphics[width=0.45\textwidth]{WCElargebasis.pdf}
    \caption{\centering Radius of convergence in the CSF basis with $n$ basis function for the MP partitioning (left) and the WC partitioning (right).}
    \label{fig:RadiusPartitioning}
\end{figure}

Différence RHF/UHF, Hamiltonien non-bloc diagonal, coefficients complexe pour R<3/2 \\

>>>>

\begin{table}[h!]
    \centering
    \caption{\centering Percentage of electron correlation energy recovered and $\expval{S^2}$ for the \ce{H2} molecule as a function of bond length (r,\si{\angstrom}) in the STO-3G basis set (taken from \cite{Gill_1988}).}
    \begin{tabular}{c c c c c c c}
\hline
 $r$ & UHF & UMP2 & UMP3 & UMP4 & $\expval{S^2}$ \\
\hline
0.75 & 0.0\% & 63.8\% & 87.4\% & 95.9\% & 0.00\\
1.35 & 0.0\% & 15.2\% & 26.1\% & 34.9\% & 0.49\\
2.00 & 0.0\% & 01.0\% & 01.8\% & 02.6\% & 0.95\\
2.50 & 0.0\% & 00.1\% & 00.3\% & 00.4\% & 0.99\\
\hline
\end{tabular}
    \label{tab:SpinContamination}
\end{table}

@@

%============================================================%
\section{The spherium model}\label{sec:spherium}
%============================================================%

@@

\begin{table}[h!]
\centering
\caption{\centering RHF and UHF energies in the minimal basis and exact energies for various R.}
\begin{tabular}{c c c c c c c c c}
$R$ & 0.1 & 1 & 2 & 3 & 5 & 10 & 100 & 1000 \\
\hline
RHF & 10 & 1 & 0.5 & 0.3333 & 0.2 & 0.1 & 0.01 & 0.001 \\
\hline
UHF & 10 & 1 & 0.490699 & 0.308532 & 0.170833 & 0.078497 & 0.007112 & 0.000703 \\
\hline
Exact & 9.783874 & 0.852781 & 0.391959 & 0.247898 & 0.139471 & 0.064525 & 0.005487 & 0.000515 \\
\end{tabular}
\label{tab:ERHFvsEUHF}
\end{table}

@@

The Hamiltonian $\bH(\lambda)$ is block diagonal because of the symmetry of the basis set ($\psi_1$ only interacts with $\psi_4$ and $\psi_2$ with $\psi_3$). The two singly excited states yields a singlet and a triplet sp\textsubscript{z} but they don't have the appropriate symmetry so these states can't form exceptional points with the ground state. However there is an avoided crossing (see Fig. \ref{fig:RHFMiniBas}) between the s\textsuperscript{2} and the p\textsubscript{z}\textsuperscript{2} which is connected to two exceptional points in the complex plane.

@@

Then using this basis set we can compare the different partitioning of \autoref{sec:AlterPart}. The \autoref{fig:RadiusPartitioning} shows the evolution of the radius of convergence $R_{\text{CV}}$ in function of $R$ for the M{\o}ller-Plesset, the Epstein-Nesbet, the Weak Correlation and the Strong Coupling partitioning in a minimal basis i.e. $P_0$ and $P_1$ and in the three function basis. We see that the radius of convergence of the strong coupling partitioning is growing with R whereas it is decreasing for the three other partitioning. This result was expected because the three decreasing partitioning use a weakly correlated reference so $\bH^{(0)}$ is a good approximation for small $R$. On the contrary the strong coupling one uses a strongly correlated reference so this series converge better when the electron are strongly correlated i.e. when $R$ is large for the spherium model.
The M{\o}ller-Plesset partitioning is always better than the weak correlation in \autoref{fig:RadiusPartitioning}. In the weak correlation partitioning the powers of $R$ are well-separated so each term of the series is a different power of $R$. Whereas the M{\o}ller-Plesset reference is proportionnal to $R^{-1}$ and $R^{-2}$ so the MP series is not well-defined in terms of powers of $R$. Moreover it can be proved that the n-th order energy of the weak correlation series can be obtained as a limited development of MP$n$ respective to $R$. It seems that the Epstein-Nesbet partitioning is better than the M{\o}ller-Plesset one for very small R in the minimal basis. In fact, it is just an artifact of the minimal basis because in the minimal basis augmented with $P_2
$ (and in larger basis set) the M{\o}ller-Plesset series converge faster for all value of $R$.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.45\textwidth]{PartitioningRCV2.pdf}
    \includegraphics[width=0.45\textwidth]{PartitioningRCV3.pdf}
    \caption{\centering Radius of convergence in the minimal basis (left) and in the minimal basis augmented with $P_2$ (right) for different partitioning of the Hamiltonian$\bH(\lambda)$.}
    \label{fig:RadiusPartitioning}
\end{figure}

The \autoref{fig:RadiusBasis} shows that the radius of convergence is not very sensitive to the expansion of the basis set. The CSF basis function have all the same spin and spatial symmetry so we expect that the singularities obtained within this basis set will be $\alpha$ singularities. The \autoref{tab:SingAlpha} proves that the singularities considered in this case are $\alpha$ singularities. This is consistent with the observation of Goodson and Sergeev \cite{Goodson_2004} on $\alpha$ singularities. The discontinuities observed in \autoref{fig:RadiusBasis} with the M{\o}ller-Plesset partitioning are due to a change of the dominant singularity. \\

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.45\textwidth]{MPlargebasis.pdf}
    \includegraphics[width=0.45\textwidth]{WCElargebasis.pdf}
    \caption{\centering Radius of convergence in the CSF basis with $n$ basis function for the MP partitioning (left) and the WC partitioning (right).}
    \label{fig:RadiusBasis}
\end{figure}

\begin{table}[h!]
\centering
\caption{\centering }
\begin{tabular}{c}
a
\end{tabular}
\label{tab:SingAlpha}
\end{table}

Now we will investigate the differences in the singularity structure between the RHF and UHF formalism. To do this we use the symmetry-broken orbitals obtained in \autoref{sec:spherium}. Thus the UHF two-electron basis is:
